The Effect of Elections and Prime Ministers on Discussion in the Australian Federal Parliament (1901--2018)

### 2019-03-31 outputs/docs/paper/paper.Rmd
model updates in paper

The topic proportions on day $d$ in chamber $c$ are modelled in reference to their membership of a particular sitting period $s$. Firstly, we assume that each distribution of topics, $\theta_{c,d,1:P}$ for each day is a draw from a Dirichlet distribution with mean parameter $\mu_{s[c,d],1:P}$:
$$
\theta_{c,d,1:P} \sim \mbox{Dirichlet}(\mu_{s[c,d],1:P})
$$
where the notation $s[c,d]$ refers to the sitting period $s$ to which day $d$ of chamber $c$ belongs. This distributional assumption accounts for the fact that on any given day in either chamber, the sum of all proportions in each topic must be one.

The goal of the model is to relate these proportions to the current prime minister $g$ and election period $e$, assuming that an 'election effect' would be at its peak in the days close to the election, then decay over the period. The mean parameters $\mu_{s,p}$ are modelled on the log scale as:

$$
\log \mu_{s,p} = \alpha_{g[c,s],p} + \beta_{e[c,s],p} \cdot (N_{s[e]} - s) + \delta_{s[c],p}
$$
where: $\alpha_{g[c,s],p}$ is the mean effect for prime ministers $g$ (which covers sitting period $s$ in chamber $c$) and topic $p$; $\beta_{e[c,s],d,p}$ is the effect of election $e$ (which occurs in sitting period $s$, in chamber $c$) for topic $p$; $N_s$ is the total number of sitting periods in election period $e$; and $\delta_{s,p}$ is a random, or levels, effect for each sitting period and topic.

The term for the prime minister, $\alpha_{g[c,s],p}$, assumes there is some underlying mean effect of each prime minister on the topic distribution. We place uninformative priors on each of these parameters:
$$
\alpha_{g[c,s],p} \sim \mbox{Normal}(0, 100).
$$

@@

Finally, the sitting-period-specific random effect $\delta_{s[c],p}$ allows for the topic distributions in some sitting periods to be different than expected based on prime minister and election effects. This allows us to identify large deviations away from the expected distribution, thus helping to identify the effect of other, non-prime-minister and non-election events. In addition, this set up also partially pools effects across sitting periods. The $\delta_{s[c],p}$ values are modelled as:
$$
\delta_{s,p} \sim \mbox{Normal}(0, \sigma_{g[c,s],p}^2).
$$

>>>>

The topic proportions on day $d$ in chamber $c$ are modelled in reference to their membership of a particular sitting period $s$. Firstly, we assume that each distribution of topics, $\theta_{c,d,1:P}$ for each day is a draw from a Dirichlet distribution with mean parameter $\mu_{c,s[d],1:P}$:
$$
\theta_{c,d,1:P} \sim \mbox{Dirichlet}(\mu_{c,s[d],1:P})
$$
where the notation $s[d]$ refers to the sitting period $s$ to which day $d$ of chamber $c$ belongs. This distributional assumption accounts for the fact that on any given day in either chamber, the sum of all proportions in each topic must be one.

The goal of the model is to relate these proportions to the current prime minister $g$ and election period $e$, assuming that an 'election effect' would be at its peak in the days close to the election, then decay over the period. The mean parameters $\mu_{c,s,p}$ are modelled on the log scale as:

$$
\log \mu_{c,s,p} = \alpha_{g[c,s],p} + \beta_{e[c,s],p} \cdot (N_{s[e]} - s) + \delta_{c,s,p}
$$
where: $\alpha_{g[c,s],p}$ is the mean effect for prime ministers $g$ (which covers sitting period $s$ in chamber $c$) and topic $p$; $\beta_{e[c,s],d,p}$ is the effect of election $e$ (which occurs in sitting period $s$, in chamber $c$) for topic $p$; $N_s$ is the total number of sitting periods in election period $e$; and $\delta_{c,s,p}$ is a random, or levels, effect for each chamber, sitting period and topic.

The term for the prime minister, $\alpha_{g[c,s],p}$, assumes there is some underlying mean effect of each prime minister on the topic distribution that is common across chambers. We place uninformative priors on each of these parameters:
$$
\alpha_{g[c,s],p} \sim \mbox{Normal}(0, 100).
$$

@@

Finally, the sitting-period-specific random effect $\delta_{c,s,p}$ allows for chamber- and sitting-period specific deviations in the topic distributions. This allows us to identify periods where there may be differences in the topics discussed across the two chambers. It also allows us to identify large deviations away from the expected distribution, thus helping to identify the effect of other, non-prime-minister and non-election events. In addition, this set up also partially pools effects across sitting periods. The $\delta_{c,s,p}$ values are modelled as:
$$
\delta_{c,s,p} \sim \mbox{Normal}(0, \sigma_{g[c,s],p}^2).
$$
